---
title: "Frass_early_spring_investigations"
output: html_document
date: "2026-10-01"
---
#load libraries
```{r}
#libraries
library(gsheet)
library(dplyr)
library(tidyr)
library(purrr)
library(rstatix)
library(tidyverse)
library(jsonlite)
library(daymetr)
library(corrplot)
library(zoo)
library(abind)
library(pastecs)
library(StatMatch)
library(MASS)
```

#loading CC data and altering it for julian week 
```{r}
#loading CC data
#Loading in CC data 
options(timeout = 300)  
api_url <- "https://api.github.com/repos/hurlbertlab/caterpillars-analysis-public/contents/data"
files <- fromJSON(api_url)
dataset_file <- files$name[grepl("fullDataset", files$name, ignore.case = TRUE)]
# pick the latest one
latest_file <- dataset_file[1]
github_raw <- "https://raw.githubusercontent.com/hurlbertlab/caterpillars-analysis-public/master/data/"
fullDataset <- read.csv(paste0(github_raw, latest_file))


#-------------------------------------------------------------------------------
#Meandensitybyweek function for caterpillar data
# Function for calculating the mode of a series of values
# --in this particular use case, if there multiple modes, we want the largest value
Mode = function(x){ 
  if (!is.numeric(x)) {
    stop("values must be numeric for mode calculation")
  }
  ta = table(x)
  tam = max(ta)
  mod = as.numeric(names(ta)[ta == tam])
  return(max(mod))
}

# Function for substituting values based on a condition using dplyr::mutate
# Modification of dplyr's mutate function that only acts on the rows meeting a condition
mutate_cond <- function(.data, condition, ..., envir = parent.frame()) {
  condition <- eval(substitute(condition), .data, envir)
  .data[condition, ] <- .data[condition, ] %>% mutate(...)
  .data
}

# Function for calculating and displaying arthropod phenology by week (density, mean biomas)
meanDensityByWeek = function(surveyData, # merged dataframe of Survey and arthropodSighting tables for a single site
                             ordersToInclude = 'All',       # which arthropod orders to calculate density for (codes)
                             
                             minLength = 0,         # minimum arthropod size to include 
                             jdRange = c(1,365),
                             outlierCount = 10000,
                             plot = FALSE,
                             plotVar = 'fracSurveys', # 'meanDensity' or 'fracSurveys' or 'meanBiomass'
                             minSurveyCoverage = 0.8, # minimum proportion of unique survey branches examined per week in order to include the week as a data point
                             allDates = TRUE,
                             new = TRUE,
                             color = 'black',
                             allCats = TRUE,
                             ...)                  

{
  
  if(length(ordersToInclude)==1 & ordersToInclude[1]=='All') {
    ordersToInclude = unique(surveyData$Group)
  }
  
  numUniqueBranches = length(unique(surveyData$PlantFK))
  
  firstFilter = surveyData %>%
    filter(julianday >= jdRange[1], julianday <= jdRange[2]) %>%
    mutate(julianweek = 7*floor(julianday/7) + 4)
  
  effortByWeek = firstFilter %>%
    group_by(julianweek) %>%
    summarize(nSurveyBranches = n_distinct(PlantFK),
              nSurveys = n_distinct(ID)) %>%
    mutate(modalBranchesSurveyed = Mode(5*ceiling(nSurveyBranches/5)),
           nSurveySets = nSurveys/modalBranchesSurveyed,
           modalSurveySets = Mode(round(nSurveySets)),
           okWeek = ifelse(nSurveySets/modalSurveySets >= minSurveyCoverage, 1, 0))
  
  if (allDates) {
    effortByWeek$okWeek = 1
  }
  
  if (!allCats) {
    secondFilter = firstFilter %>%
      filter(Hairy != 1, Tented != 1, Rolled != 1)
  } else {
    secondFilter = firstFilter
  }
  
  arthCount = secondFilter %>%
    filter(Length >= minLength, 
           Group %in% ordersToInclude) %>%
    mutate(Quantity2 = ifelse(Quantity > outlierCount, 1, Quantity)) %>% #outlier counts replaced with 1
    group_by(julianweek) %>%
    summarize(totalCount = sum(Quantity2, na.rm = TRUE),
              numSurveysGTzero = length(unique(ID[Quantity > 0])),
              totalBiomass = sum(Biomass_mg, na.rm = TRUE)) %>% 
    right_join(effortByWeek, by = 'julianweek') %>%
    filter(okWeek == 1) %>%
    #next line replaces 3 fields with 0 if the totalCount is NA
    mutate_cond(is.na(totalCount), totalCount = 0, numSurveysGTzero = 0, totalBiomass = 0) %>%
    mutate(meanDensity = totalCount/nSurveys,
           fracSurveys = 100*numSurveysGTzero/nSurveys,
           meanBiomass = totalBiomass/nSurveys) %>%
    arrange(julianweek) %>%
    data.frame()
  
  if (plot & new) {
    plot(arthCount$julianweek, arthCount[, plotVar], type = 'l', 
         col = color, las = 1, ...)
    points(arthCount$julianweek, arthCount[, plotVar], pch = 16, col = color, ...)
  } else if (plot & new==F) {
    points(arthCount$julianweek, arthCount[, plotVar], type = 'l', col = color, ...)
    points(arthCount$julianweek, arthCount[, plotVar], pch = 16, col = color, ...)
  }
  return(arthCount)
}
#-------------------------------------------------------------------------------
#site filter fulldataset for all years
cat_data_all_years <-fullDataset %>%
  filter(Name %in% c("NC Botanical Garden", "Prairie Ridge Ecostation"),
         Year %in% 2015:2026) 

#have meandensitybyweek aggregate caterpillar stuff by week 
cat_data_byweek <- cat_data_all_years %>%
  group_by(Year, Name) %>%
  group_split() %>%                 # split into a list, one dataframe per year
  map_dfr(~ {
    out <- meanDensityByWeek(
      surveyData = .x,
      ordersToInclude = "caterpillar",
      allDates = TRUE
    )
    out$Year <- unique(.x$Year)      # add Year back
    out$Name <- unique(.x$Name)
    out
  })%>%
  rename(Site=Name)%>%
  mutate(Site = case_when(
    Site == "NC Botanical Garden" ~ "117",
    Site == "Prairie Ridge Ecostation" ~ "8892356"  )) 
#with cat_data_byweek want to use fracSurvey and meanDensity columns


#filter by cutoff days and correct years for both sites
cat_data_byweek <- cat_data_byweek %>%
  filter(
    (Site == 117 & Year %in% c(2015, 2018, 2019, 2021, 2022) & julianweek %in% 142:200) |
      (Site == 8892356 & Year %in% c(2015:2019, 2021:2026) & julianweek %in% 154:198))


```

#   Reading in Frass data  
```{r}
# Function for reading in frass data from GoogleDoc
# *if aim is to backup GoogleDoc and write to disk only, then open =F and write = T
# *if aim is to use data without writing to disk, then open = T and write = F
frassData = function(open = T, write = F) {
  require(gsheet)
  url = "https://docs.google.com/spreadsheets/d/1RwXzwhHUbP0m5gKSOVhnKZbS1C_NrbdfHLglIVCzyFc/edit#gid=1479231778"
  data = gsheet2tbl(url)
  
  if (write) {
    # Write a copy
    write.csv(data, paste('data/frass_', Sys.Date(), '.csv', sep = ''),
              row.names = F)
  }
  if (open) { return (data) }
}
#----------------------------------------------------------------------------------
#Function for fixing time format and downloading corrected csv
TimeCleaning = function() {
  read_in_data <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1RwXzwhHUbP0m5gKSOVhnKZbS1C_NrbdfHLglIVCzyFc/edit#gid=1479231778')
  
  remove_NAs <- read_in_data %>%
    filter(!is.na(Time.Set) & !is.na(Time.Collected))
  
  write.csv(remove_NAs %>% 
              mutate(Time.Set = ifelse(test = grepl(":", remove_NAs$Time.Set), 
                                       yes = remove_NAs$Time.Set, 
                                       no = paste(substr(remove_NAs$Time.Set, 1, nchar(remove_NAs$Time.Set)-2), ":", substr(remove_NAs$Time.Set, 3, 4), sep = "")), 
                     Time.Collected = ifelse(test = grepl(":", remove_NAs$Time.Collected), 
                                             yes = remove_NAs$Time.Collected, 
                                             no = paste(substr(remove_NAs$Time.Collected, 1, nchar(remove_NAs$Time.Collected)-2), ":", substr(remove_NAs$Time.Collected, 3, 4), sep = ""))), 
            paste('data/frass_', Sys.Date(), '.csv', sep = ''), row.names = F)
}
#----------------------------------------------------------------------------------
# Function that takes a date field (formatted as %m/%d/%Y) and a time field (hh:mm in 24h time), converts the date to julian day and adds the fractional
julianDayTime = function(date, hour_min) {
  require(lubridate)
  jday = yday(date)
  temp = sapply(strsplit(hour_min, ":"), function(x) { #(day represented by the hours and minutes)
    x = as.numeric(x)
    x[1] + x[2]/60
  })
  output = jday + temp/24
  return(output)
}
#-----------------------------------------------------------------------------------------------
# altering frassdata so that times and days are corrected
data = frassData(open = T) %>%
  filter(!is.na(Time.Set) & !is.na(Time.Collected)) %>%
  mutate(Date.Set = as.Date(Date.Set, format = "%m/%d/%Y"),
         Time.Set = as.character(Time.Set),
         Time.Collected = as.character(Time.Collected),
         Date.Collected = as.Date(Date.Collected, format = "%m/%d/%Y"),
         Year = format(Date.Collected, "%Y"),
         jday.Set = julianDayTime(Date.Set, Time.Set),
         jday.Collected = julianDayTime(Date.Collected, Time.Collected),
         frass.mg.d = Frass.mass..mg./(jday.Collected - jday.Set),
         frass.no.d = Frass.number/(jday.Collected - jday.Set),
         jday = (floor(jday.Collected) + floor(jday.Set))/2)


```

#creating occurrence frass data
```{r}

#-----------------------------------------------------------------------------------------------
#Filtering Data for Frass OCCURRENCE -> use occurance_frass_combined_weeks DF
#-----------------------------------------------------------------------------------------------
#filter data so only reliable rows are left then filter frass.mg.d so that only traps with total mass >4mg are left
filtered_mass <- data %>%
  filter(OK==1)%>% #only days deemed reliable left
  mutate(julianweek = 7 * floor(jday / 7) + 4)%>%
  mutate(included_in_trap_count = if_else(Frass.mass..mg. > 4,1,0))  #CHANGE THRESHOLD HERE

#-----------------------------------------------------------------------------------------------
#now need to count number of traps with same julian day  
occurance_frass <- filtered_mass %>%
  group_by(Site, Year, julianweek)%>%
  mutate(trap_occurance_percent = mean(included_in_trap_count == 1, na.rm=TRUE)) #mean here calculates the proportion along the length of the groupby columns


#issue that in 2015-2023 sites sampled twice a week, so still percentage works since dividing by total seen???
occurance_frass_combined_weeks <- occurance_frass %>%
  group_by(Site, Year, julianweek) %>%
  summarise(
    # average frass measurements
    trap_occurance_percent = mean(trap_occurance_percent, na.rm = TRUE),
    # keep representative values for the rest
    jday = min(jday, na.rm = TRUE),
    .groups = "drop")%>%
  mutate(Year=as.integer(Year)) %>%
  mutate(Site = case_when(
    Site == "Botanical Garden" ~ "8892356",
    Site == "Prairie Ridge" ~ "117"  )) %>%
  filter(
    (Site == 117 & Year %in% c(2015, 2018, 2019, 2021, 2022) & julianweek %in% 142:200) |
      (Site == 8892356 & Year %in% c(2015:2019, 2021:2026) & julianweek %in% 154:198))


```

#Frass number of pellets
```{r}

#-----------------------------------------------------------------------------------------------
#Filtering Data for Frass NUMBER of PELLETS 
#-----------------------------------------------------------------------------------------------
frass_pellet_number <- data %>%
  filter(OK==1)%>% #only days deemed reliable left
  mutate(julianweek = 7 * floor(jday / 7) + 4)%>%
  group_by(julianweek, Site, Year) %>%
  summarize(number_pellets = sum(Frass.number), .groups = "drop") %>%
  mutate(Site = recode(Site, "Botanical Garden" = "8892356"))%>%
  mutate(Site = recode(Site, "Prairie Ridge" = "117"))
frass_pellet_number$Year <- as.integer(frass_pellet_number$Year)



```

#Filtering Data for Frass MASS and cateprillar biomass -> IMPUTATION use imputation_data DF
```{r}
# using data to find mean frass per day for reliable frass only 
#read in proper url, change dates and label events for below meanfrass
url = "https://docs.google.com/spreadsheets/d/1RwXzwhHUbP0m5gKSOVhnKZbS1C_NrbdfHLglIVCzyFc/edit#gid=1611171427"
events = gsheet2tbl(url)
events$date = as.Date(events$date, format = "%m/%d/%Y")

#fiter out OK days and creating by trap area for mass values
meanfrass = data %>%
  filter(!is.na(Frass.mass..mg.)) %>%
  filter(OK == 1) %>% #keeps reliable frass row
  mutate(site = as.character(ifelse(Site=="Botanical Garden", 8892356, 117))) %>%
  group_by(site, Date.Collected, Year, jday) %>%
  summarize(
    mass = mean(frass.mg.d, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate( #altering the density based on size of the trap
    trap_area_cm2 = ifelse(Year <= 2018, 309.74, 197.71), #before 2018 use 309cm^2 after use 197cm^2 
    density_mg_cm2 = mass / trap_area_cm2,
    density_mg_m2 = density_mg_cm2 * 10000   # optional but recommended
  ) %>%
  left_join(events[, c('date', 'site', 'reliability')],
            by = c('Date.Collected' = 'date', 'site' = 'site')) %>%
  rename(date = Date.Collected)

#make sure frass has correct columns--------------------------------------
#having mean Frass data sorted by julian week
meanfrass <- meanfrass %>%
  mutate(julianweek = 7 * floor(jday / 7) + 4)%>% #this is using jday calculated from 'data' where it uses day and time set to alter jday
  mutate(Year = as.integer(Year)) #make sure integer
#combine and average meanfrass data that comes from same week (2015-2023 this happened), is mean frass per day
meanfrass_combinedweeks <- meanfrass %>%
  group_by(site, Year, julianweek) %>%
  summarise(
    # average frass measurements
    mass = mean(mass, na.rm = TRUE),
    density = mean(density_mg_cm2, na.rm = TRUE),
    # keep representative values for the rest
    date = min(date, na.rm = TRUE),   # or first(date)
    jday = mean(jday, na.rm = TRUE),
    reliability = first(reliability),
    
    .groups = "drop")

#Caterpillar Count data---------------------------------------------------
#NCBG site filter fulldataset for all years
NCBG <- fullDataset %>%
  filter(Name %in% c("NC Botanical Garden"),
         Year %in% 2015:2026)
#PR site filter fulldataset for all years
PR <- fullDataset %>%
  filter(Name %in% c("Prairie Ridge Ecostation"),
         Year %in% 2015:2026)
#have meandensitybyweek aggregate caterpillar stuff by week for NCBG
cats_NCBG <- NCBG %>%
  group_by(Year) %>%
  group_split() %>%                 # split into a list, one dataframe per year
  map_dfr(~ {
    out <- meanDensityByWeek(
      surveyData = .x,
      ordersToInclude = "caterpillar",
      allDates = TRUE
    )
    out$Year <- unique(.x$Year)      # add Year back
    out
  })%>%
  mutate(site=8892356)
#have meandensitybyweek aggregate caterpillar stuff by week for PR
cats_PR <- PR %>%
  group_by(Year) %>%
  group_split() %>%                 # split into a list, one dataframe per year
  map_dfr(~ {
    out <- meanDensityByWeek(
      surveyData = .x,
      ordersToInclude = "caterpillar",
      allDates = TRUE
    )
    out$Year <- unique(.x$Year)      # add Year back
    out
  })%>%
  mutate(site=117)
#all caterpillar data together
cats_all <- rbind(cats_NCBG, cats_PR)

#combine into one df
all_data <- cats_all %>%
  mutate(site = as.character(site)) %>% #make sure same type to join
  full_join(meanfrass_combinedweeks, by = c("julianweek", "Year", "site")) #shows where missing frass data is? is full join appropriate here?

#filter by cutoff days and correct years for both sites
all_data <- all_data %>%
  filter(
    (site == 117 & Year %in% c(2015, 2018, 2019, 2021, 2022) & julianweek %in% 142:200) |
      (site == 8892356 & Year %in% c(2015:2019, 2021:2026) & julianweek %in% 154:198)
  ) #ok kinda shows weeks where no frass data compared to CC but doesnt address issues of individual days where no data

#clean all data so only have columns I want and divide by trap area
all_data_clean <- all_data %>%
  dplyr::select(site, Year, jday, julianweek, meanBiomass, date, mass) %>% #make sure no duplicate weeks
  group_by(site, Year, julianweek) %>%
  summarise(
    meanBiomass = mean(meanBiomass, na.rm = TRUE),
    mass        = mean(mass, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(biomass_density = meanBiomass/(ifelse(Year <= 2018, 309.74, 197.71))) %>% #dividing by 209 for years 2018 and before, and 197 for years after
  mutate(frass_density = mass/ (ifelse(Year <= 2018, 309.74, 197.71)))


#bring in imputation data from last semester:
#imputation function
impute_biomass_data <- function(data, site_mass_defaults) {
  
  # --- Precompute first/last julianweek reference values across years ---
  first_jweek_refs <- data %>%
    group_by(site, Year) %>%
    slice_min(julianweek, n = 1) %>%
    ungroup() %>%
    group_by(site) %>%
    summarise(
      ref_first_meanBiomass = mean(meanBiomass, na.rm = TRUE),
      ref_first_mass        = mean(mass, na.rm = TRUE),
      .groups = "drop"
    )
  
  last_jweek_refs <- data %>%
    group_by(site, Year) %>%
    slice_max(julianweek, n = 1) %>%
    ungroup() %>%
    group_by(site) %>%
    summarise(
      ref_last_meanBiomass = mean(meanBiomass, na.rm = TRUE),
      ref_last_mass        = mean(mass, na.rm = TRUE),
      .groups = "drop"
    )
  
  data %>%
    left_join(first_jweek_refs, by = "site") %>%
    left_join(last_jweek_refs,  by = "site") %>%
    group_by(site, Year) %>%
    arrange(julianweek, .by_group = TRUE) %>%
    mutate(
      # --- save originals before imputation ---
      orig_meanBiomass = meanBiomass,
      orig_mass        = mass,
      
      # --- fill first julianweek NA with cross-year average for that site ---
      meanBiomass = if_else(
        is.na(meanBiomass) & julianweek == first(julianweek),
        ref_first_meanBiomass,
        meanBiomass
      ),
      mass = if_else(
        is.na(mass) & julianweek == first(julianweek),
        coalesce(site_mass_defaults[as.character(site)], ref_first_mass),
        mass
      ),
      
      # --- fill last julianweek NA with cross-year average for that site ---
      meanBiomass = if_else(
        is.na(meanBiomass) & julianweek == last(julianweek),
        ref_last_meanBiomass,
        meanBiomass
      ),
      mass = if_else(
        is.na(mass) & julianweek == last(julianweek),
        ref_last_mass,
        mass
      ),
      
      # --- interpolate interior NAs ---
      meanBiomass = na.approx(meanBiomass, x = julianweek, na.rm = FALSE, rule = 2, maxgap = 2),
      mass        = na.approx(mass,        x = julianweek, na.rm = FALSE, rule = 2, maxgap = 2),
      
      # --- flag imputed rows ---
      imputed_biomass = as.integer(is.na(orig_meanBiomass) & !is.na(meanBiomass)),
      imputed_mass    = as.integer(is.na(orig_mass)        & !is.na(mass)),
      
      # --- drop helper columns ---
      orig_meanBiomass      = NULL,
      orig_mass             = NULL,
      ref_first_meanBiomass = NULL,
      ref_first_mass        = NULL,
      ref_last_meanBiomass  = NULL,
      ref_last_mass         = NULL
    ) %>%
    ungroup()
}
#run function
imputation_data <- impute_biomass_data(all_data_clean, site_mass_defaults = c() )
#make site become Site so can combine later
imputation_data <- imputation_data %>%
  rename(Site = site)


```

#combining frass mass/occurrence and caterpillar occurrence,biomass, and density into one dataframe: 
```{r}
all_six_variables_dataframe <- cat_data_byweek %>%
  dplyr::select(Site, Year, julianweek, fracSurveys, meanDensity)%>%
  left_join(occurance_frass_combined_weeks, by=c("Site", "Year", "julianweek")) %>%
  left_join(imputation_data %>% dplyr::select(Site, Year, julianweek, meanBiomass, mass), by=c("Site", "Year", "julianweek")) %>%
  left_join(frass_pellet_number, dplyr::select(Site, Year, julianweek, number_pellets), by =c("Site", "Year", "julianweek")) %>%
  rename(frass_mass = mass) #all years previous standardized (same jday range before joining)

```

#calculate centroid dates and total june values for all 6 variables
```{r}
vars_of_interest <- c("fracSurveys", "meanBiomass", "meanDensity", "trap_occurance_percent", "frass_mass", "number_pellets")

two_metric_variables <- all_six_variables_dataframe %>%
  group_by(Year, Site)%>%
  summarise(centroid_frass_occurrence = weighted.mean(julianweek, trap_occurance_percent, na.rm = TRUE),
         centroid_frass_mass = weighted.mean(julianweek, frass_mass, na.rm = TRUE),
         centroid_frass_pelletnum = weighted.mean(julianweek, number_pellets, na.rm = TRUE),
         centroid_cat_occurrence = weighted.mean(julianweek, fracSurveys, na.rm = TRUE),
         centroid_cat_biomass = weighted.mean(julianweek, meanBiomass, na.rm = TRUE),
         centroid_cat_density = weighted.mean(julianweek, meanDensity, na.rm = TRUE),
         total_frass_occurrence_june = sum(trap_occurance_percent[julianweek >= 152 & julianweek <= 181], na.rm = TRUE),
         total_frass_mass_june = sum(frass_mass[julianweek >= 152 & julianweek <= 181], na.rm = TRUE),
         total_frass_pelletnum_june = sum(number_pellets[julianweek >= 152 & julianweek <= 181], na.rm = TRUE),
         total_cat_occurrence_june = sum(fracSurveys[julianweek >= 152 & julianweek <= 181], na.rm = TRUE),
         total_cat_biomass_june = sum(meanBiomass[julianweek >= 152 & julianweek <= 181], na.rm = TRUE),
         total_cat_density_june = sum(meanDensity[julianweek >= 152 & julianweek <= 181], na.rm = TRUE))

```

#read in temp data and alter it for april and may mean averages
```{r}
#loading in temp data (from Nosa)------------------------------------------------------
tmp_file = tempfile(fileext = ".csv")

AnomalySites = fullDataset[, c("Name","Year", "Latitude", "Longitude")] %>% 
  filter(Name %in% c("NC Botanical Garden", "Prairie Ridge Ecostation")) %>% 
  group_by(Name, Latitude, Longitude) %>% 
  summarise(n = n()) %>% 
  dplyr::select(-n) %>% as.data.frame()   

AnomalyDaymetr = AnomalySites %>% 
  rename(
    site = Name,
    lat = Latitude,
    lon = Longitude
  ) %>%
  write.csv(tmp_file, row.names = FALSE)

# pass temp CSV to function
TempAnomaly = download_daymet_batch(
  file_location = tmp_file,
  start = 2015,
  end = 2025, # this is the most recent available in daymetr
  internal = TRUE
)
# remove temporary file 
unlink(tmp_file)

TempAnomalyData_clean <- lapply(TempAnomaly, function(x) {
  x$data %>% mutate(site = x$site,
                    Latidue = x$latitude,
                    Longitude = x$longitude)})

AllTemp= bind_rows(TempAnomalyData_clean) #this is the file we want (has all NCBG and PR data)
#clean up globals
rm(AnomalySites, TempAnomaly, files, TempAnomalyData_clean)

# *+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+
#   altering temperature to be what sites I want and have jday and jweek columns
# *+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+*+ 
#rename site names so match other dataframes, also rename so can be joined properly with meanfrass data
AllTemp <- AllTemp %>%
  mutate(site = case_when(
    site == "NC Botanical Garden" ~ "117",
    site == "Prairie Ridge Ecostation" ~ "8892356"  )) 
AllTemp <- rename(AllTemp, jday=yday)
AllTemp<- rename(AllTemp, Year=year)
AllTemp <- rename (AllTemp, Site=site)


#adding average weekly temp column to alltemp data, and averaging temp for all days with same jweek value
AllTemp_weekly <- AllTemp %>%
  mutate(julianweek = 7 * floor(jday / 7) + 4) %>%
  mutate(avgtemp = (tmax..deg.c. + tmin..deg.c.) / 2)%>%
  group_by(Year, Site, julianweek) %>%
  summarise(weeklytemp = mean(avgtemp, na.rm = TRUE),
            .groups = "drop")

#calculate avg temp for april and may combined
alltemp_weekly_spring_only <- AllTemp_weekly %>%
  group_by(Year, Site) %>%
  summarise(spring_avg_both = mean(weeklytemp[julianweek >= 91 & julianweek <=151]),
            spring_avg_may = mean(weeklytemp[julianweek >=121 & julianweek <=151]))

```

#join centroid/totals for 6 variables and avg spring temp data into one df
```{r}
temp_and_metric_combined <- two_metric_variables %>%
  left_join(alltemp_weekly_spring_only, by=c("Site","Year")) #2026 data for temp not listed yet



```

#run correlations between each metric(june totals vs centroid dates), and temperature and see what correlaiton is over time
```{r}
centroid_metric <- c("centroid_frass_occurrence", "centroid_frass_mass", "centroid_frass_pelletnum", "centroid_cat_occurrence", "centroid_cat_biomass", "centroid_cat_density")

junetotal_metric <- c("total_frass_occurrence_june", "total_frass_mass_june", "total_frass_pelletnum_june", "total_cat_occurrence_june", "total_cat_biomass_june", "total_cat_density_june")

#Pearsons (keep safe ones (non NA) only)
safe_cor_mat <- possibly(
  ~ rstatix::cor_mat(.x, method = "pearson"),
  otherwise = NULL
)
#pearsons
nested_data_pearson <- temp_and_metric_combined %>% #change DF if needed
  dplyr::select(Site, Year, all_of(centroid_metric)) %>%
  group_by(Site, Year) %>%
  nest() %>%
  mutate(cor_matrix = map(data, safe_cor_mat))

#Spearmans
safe_cor_mat_spearman <- possibly(
  ~ rstatix::cor_mat(.x, method = "spearman"),
  otherwise = NULL
)

nested_data_spearman <- temp_and_metric_combined %>%
  dplyr::select(Site, Year, all_of(centroid_metric)) %>%
  group_by(Site, Year) %>%
  nest() %>%
  mutate(cor_matrix = map(data, safe_cor_mat_spearman))

#---------------------------------------------------------------------------
#visualizations- may have to do for loop
correlation_plotting <- function(data, year_choice, site_choice) {
  df <- data %>%
    filter(Year == year_choice, Site == site_choice)
  #pull the cor_matrix out of the list-column
  cor_df <- df$cor_matrix[[1]]
  
  #convert rstatix cor_mat() output (has a rowname/var column) into a real matrix
  cor_matrix <- cor_df %>%
    column_to_rownames(var = colnames(cor_df)[1]) %>%
    as.matrix()
  ## plot:
  corrplot(cor_matrix, 
           type = "upper", 
           title = paste(site_choice, year_choice),
           mar = c(0, 0, 2, 0),
           method = "shade", 
           order = "original", #original,hclust, alphabet
           tl.col = "black", 
           tl.srt = 45,
           cl.align.text="l",
           cl.offset = .5,
           addCoef.col = "black",   
           number.cex = 0.8,
           addgrid.col = "black",
           col = colorRampPalette(c("firebrick2", "white", "dodgerblue3"))(200))
  
  invisible(cor_matrix)
}
correlation_plotting(nested_data_pearson, 2022, 8892356)  

##saving as a pdf------------------------ ^^^^^
# Years for each site
years_PR   <- c(2015, 2018, 2019, 2021, 2022)
years_NCBG <- setdiff(2015:2026, 2020)   
setwd("C:/Z_School/school/HurlbertLab/graphs")
#set up pdf
pdf(
  file = "7vars_correlation_pearson.pdf",
  width = 8,
  height = 8)
#layout for pdf
par(
  mfrow = c(3, 2),
  mar = c(4, 4, 3, 6),  
  oma = c(0, 0, 2, 0))
#loops over each sites
for (yr in years_NCBG) {
  try(
    correlation_plotting(
      data = nested_data_pearson,
      year_choice = yr,
      site_choice = 8892356  
    ),
    silent = TRUE)}
for (yr in years_PR) {
  try(
    correlation_plotting(
      data = nested_data_pearson,
      year_choice = yr,
      site_choice = 117   
    ),
    silent = TRUE)}
dev.off()


```


















#linear model with site as covariant
```{r}
library(dplyr); library(purrr); library(broom)

dat <- temp_and_metric_combined %>%
  filter(!is.na(spring_avg_both)) %>%
  mutate(Site = factor(Site),
         temp_z = scale(spring_avg_both)[, 1],
         across(starts_with("total_"), log1p, .names = "{.col}_log"))

responses <- c("centroid_frass_occurrence","centroid_frass_mass","centroid_frass_pelletnum",
               "centroid_cat_occurrence","centroid_cat_biomass","centroid_cat_density",
               "total_frass_occurrence_june_log","total_frass_mass_june_log",
               "total_frass_pelletnum_june_log","total_cat_occurrence_june_log",
               "total_cat_biomass_june_log","total_cat_density_june_log")

results <- map_df(responses, function(y) {
  d <- dat %>% filter(!is.na(.data[[y]]))
  m <- lm(reformulate(c("temp_z", "Site"), y), data = d)
  ci <- confint(m)["temp_z", ]
  tidy(m) %>% filter(term == "temp_z") %>%
    mutate(response = y, n = nrow(d),
           lwr = ci[1], upr = ci[2],
           partial_R2 = summary(m)$r.squared -
                        summary(lm(reformulate("Site", y), data = d))$r.squared)
}) %>%
  mutate(p_adj = p.adjust(p.value, method = "BH"))

#interpret:
#Slope for centroids is days shifted per 1 SD of spring temperature, which is the most interpretable result you'll get.
#Partial R² is the extent to which temperature explains variation beyond site differences.
#Confidence intervals matter more than p-values at this sample size.


```






