---
title: "Analysis of frass data and frass methodology at UNC"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r loading script file, include=FALSE}
source("C:/git/frass/analysis.r")
```
========================================================================================================================

These data are being collected as part of the citizen science project *Caterpillars Count!* based in the Hurlbert lab at the University of North Carolina.

This project is testing frass trap prototypes using repurposed milk jugs, with a broader goal to create a streamlined and accessible frass collection method for citizen scientists. Currently, the data are being analysized to evaluate the reliability of this collection method, and of image-based estimates for frass weight.

**Repository location**: https://github.com/hurlbertlab/frass
**Image processing package**: https://fiji.sc/

Here's a plot comparing the weight of the "raw" frass collection (i.e. debris included) vs. the "sorted" frass (in which trained assistants have removed all non-frass pieces). All plots have been normalized for outliers in the data.

```{r echo=FALSE, results='hide',message=FALSE}
raw_sort_outlier_excl = lm(Weight_Sorted ~ Weight_Raw, data = dataWO)
plot(data$Weight_Raw[data$Weight_Raw<50], data$Weight_Sorted[data$Weight_Raw<50],main = "Frass Weight Comparison (mg.)", xlab = "Weight Raw", ylab = "Weight Sorted", pch = 17, cex = 1, col = 'red')
abline(raw_sort_outlier_excl)
sortraw_sum = summary(raw_sort_outlier_excl)
sortraw_sum_r2 = sortraw_sum$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(sortraw_sum_r2, digits = 3)))
text(x = 3.3, y = 16.2, labels = mylabel)
```

And here is the relationship between the raw frass pieces collected and the percent of area the raw frass pieces occupy using a standard backdrop, estimated by the image processer, Fiji.

```{r echo=FALSE, results='hide',message=FALSE}
plot(data$Pieces_Sorted[data$Pieces_Raw<60], data$Img_Raw[data$Pieces_Raw<60], main = "Raw Frass Comparison: # of Pieces vs. % of Area", 
     xlab = "Total Pieces", ylab = "% of Area (unsorted)", pch = 20, cex = 1, col = 'orange')
raw_pcs = lm(Img_Raw ~ Pieces_Sorted, data = data)
raw_pcs_outlier_excl = lm(Img_Raw ~ Pieces_Sorted, data = data_rawpcsWO)
abline(raw_pcs_outlier_excl)
sortrawimg_sum = summary(raw_pcs_outlier_excl)
sortrawimg_sum_r2 = sortrawimg_sum$adj.r.squared
mylabel = bquote(italic(R)^2 == .(format(sortrawimg_sum_r2, digits = 3)))
text(x = 4.9, y = 6.1, labels = mylabel)
```

