---
title: "Analysis of frass data and frass methodology at UNC"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


========================================================================================================================

These data are being collected as part of the citizen science project *Caterpillars Count!* based in the Hurlbert lab at the University of North Carolina.

This project is testing frass trap prototypes using milk jugs, with a broader goal to create a streamlined and accessible frass collection method for for citizen scientists. Currently, the data are being analysized to evaluate the reliability of this collection method, and of image-based estimates for frass weight. 

https://github.com/hurlbertlab/frass

Here's a plot comparing the weight of the "raw" frass collection (i.e. debris included) vs. the "sorted" frass (in which trained assistants have removed the debris)

```{r loading script file, include=FALSE}
source("C:/git/frass/analysis.r")
```

```{r echo=FALSE, results='hide',message=FALSE}
plot(data$Pieces_Raw[data$Pieces_Raw<60], data$Img_Raw[data$Pieces_Raw<60], main = "Raw Frass Comparison: # of Pieces vs. % of Area", 
     xlab = "Total Pieces (frass w/ debris)", ylab = "% of Area", pch = 20, cex = 1, col = 'orange')
raw_pcs = lm(Img_Raw ~ Pieces_Raw, data = data)
raw_pcs_outlier_excl = lm(Img_Raw ~ Pieces_Raw, data = data_rawpcsWO)
abline(raw_pcs_outlier_excl)
summary(raw_pcs_outlier_excl)
```